Dac summary 01/19/2021

### 2021-11-13 final_dac_summary_pull_20210119.Rmd
prelim analyses

---
title: "Dac summary 01/19/2021"
author: "Erica Baller"
date: "4/15/2021"
output: html_document
---

@@

Date of FINAL Pull: 1/19/2021
Spreadsheet: /Users/eballer/BBL/msdepression/data/dac/investigatingdepressioninmspatients.csv 
Summary of request to Victoria @ DAC:
---
9/2/2020

@@

This script goes through data that has been returned from the DAC on 1/19 and summarizes it
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

@@

#from Jan 2021 pull
data <- read.csv("/Users/eballer/BBL/msdepression/data/dac/investigatingdepressioninmspatients.csv", header = TRUE, stringsAsFactors = FALSE, na.strings = c("NULL"), sep =",")

@@

@@

@@

@@

@@

@@

#sex
hist(data_empi_acc_f_phq$sex_binarized[which(!is.na(data_empi_acc_f_phq$sex_binarized))], 
     main = paste0("Sex Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$sex_binarized))), "; M = ", length(which(data_empi_acc_f_phq$sex_binarized == 1)), "; F = ", length(which(data_empi_acc_f_phq$sex_binarized == 2)) ), 
     xlab = "Male = 1, Female = 2", 
     breaks = 2)

@@

```
```{r frequencies_of_dep_in_phq2_vs_9}
num_dep_phq2 <- length(which(!is.na(data_empi_acc_f_phq$PHQ.2) & (data_empi_acc_f_phq$PHQ.2 >= 3))) #n = 130
num_dep_phq9 <- length(which(!is.na(data_empi_acc_f_phq$PHQ.9) & (data_empi_acc_f_phq$PHQ.9 >= 5))) #n = 184; https://www.sciencedirect.com/science/article/pii/B9780128179338000293 recommends using cut point of 11
num_total_phq2 <-length(which(!is.na(data_empi_acc_f_phq$PHQ.2)))
num_total_phq9 <-length(which(!is.na(data_empi_acc_f_phq$PHQ.9))) 
dep_total_phq_table <- data.frame(matrix(data = c(num_dep_phq2, num_dep_phq9, num_total_phq2, num_total_phq9), nrow = 2, ncol = 2))
names(dep_total_phq_table) <- c("dep", "healthy")
row.names(dep_total_phq_table) <- c("phq2", "phq9")
chisq_dep_phq2_v_phq9 <- chisq.test(dep_total_phq_table)
chisq_dep_phq2_v_phq9
print("observed")
chisq_dep_phq2_v_phq9$observed
print("expected")
chisq_dep_phq2_v_phq9$expected
print("residuals")
chisq_dep_phq2_v_phq9$residuals
```

@@

```
```{r comparing_out_group_with_Tim_group}
##### checking to see whether we have clinical data for patients who have already been segmented as of 5/4/2021
### had to do some parsing of tim's spreadsheet: /Users/eballer/BBL/msdepression/data/insurance_images_for_depression.csv
### used the following commands to extract the empi and the date -> in tim's they are in the format sub-***** and ses-YYYYMMDD

@@

segmented <- read.csv("/Users/eballer/BBL/msdepression/data/insurance_images_for_depression_with_accession.csv", header = F, sep = ",") #n = 4468
names(segmented) <- c("EMPI", "EXAM_DATE", "SES", "ACCESSION_NUM")

@@

@@

print(paste0("N with ICD 9/10 codes for depression and unique EMPI = ", num_unique_EMPI_and_dep, " out of ", dim(overlap_group)[1]))

@@

>>>>

---
title: "Dac summary 01/19/2021"
author: "Erica Baller"
date: "9/29/2021"
output:
  pdf_document: default
  html_document: default
---

@@

Date of FINAL Pull: 1/19/2021
Spreadsheet: /Users/eballer/BBL/msdepression/data/dac/investigatingdepressioninmspatients_dates_right_format.csv 
Summary of request to Victoria @ DAC:
---
9/2/2020

@@

This script goes through data that has been returned from the DAC on 1/19 and summarizes it
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(visreg)
library(lme4)
library(lmerTest)
#functions
source("~/BBL/msdepression/scripts/msdepression_functions.R")
```

@@

#from Jan 2021 pull
data <- read.csv("/Users/eballer/BBL/msdepression/data/dac/investigatingdepressioninmspatients_dates_right_format.csv", header = TRUE, stringsAsFactors = FALSE, na.strings = c("NULL"), sep =",")

@@

#make accession # and lab findings into integer type, contained in row 28:42, Margin=2 to work on columns
data_empi_acc_f_phq$ACCESSION_NUM <- as.integer(data_empi_acc_f_phq$ACCESSION_NUM)
data_empi_acc_f_phq[27:29] <- apply(X = data_empi_acc_f_phq[27:29], FUN = as.integer, MARGIN = 2)
data_empi_acc_f_phq[31:45] <- apply(X = data_empi_acc_f_phq[31:45], FUN = as.integer, MARGIN = 2)
data_empi_acc_f_phq$sex_binarized <- ifelse(data_empi_acc_f_phq$SEX == "MALE", 1, 2)
data_empi_acc_f_phq$race_binarized <- ifelse(data_empi_acc_f_phq$RACE == "WHITE", 1, 2)
#get date into better format
data_empi_acc_f_phq$EXAM_DATE <- as.Date(data_empi_acc_f_phq$BEGIN_EXAM_DTTM, format = "%m/%d/%y")
data_empi_acc_f_phq$EXAM_DATE <- as.numeric(gsub(data_empi_acc_f_phq$EXAM_DATE, pattern = "-", replacement = ""))
########################################
#### Start counting / making subsets ###
########################################

@@

print(paste0("Mean dep + PHQ-2 = ", round(mean(!is.na(dep_df$PHQ.2)),3), "[",round(sd(!is.na(dep_df$PHQ.2)),3), "]; n = ", length(which(dep_df$PHQ.2 >= 3)), " meeting criteria for depression (PHQ2) >= 3")) #n = 79

@@

print(paste0("Mean dep + PHQ-9 = ", round(mean(!is.na(dep_df$PHQ.9)),3), "[",round(sd(!is.na(dep_df$PHQ.9)),3), "]; n = ", length(which(dep_df$PHQ.9 >= 10)), " meeting criteria for depression (PHQ9) >= 10")) #n = 79

@@

print(paste0("Mean/SD healthy w/PHQ-2 = ", round(mean(!is.na(healthy_df$PHQ.2)),3), "[",round(sd(!is.na(healthy_df$PHQ.2)),3), "]; n = ", length(which(healthy_df$PHQ.2>= 3)), " meeting criteria for depression (PHQ2) >= 3"))

@@

print(paste0("Mean/SD healthy w/PHQ-9 = ", round(mean(!is.na(healthy_df$PHQ.9)),3), "[",round(sd(!is.na(healthy_df$PHQ.9)),3), "]; n = ", length(which(healthy_df$PHQ.9>= 10)), " meeting criteria for depression (PHQ9) >= 10"))

#statistical testing
t.test(!is.na(healthy_df$PHQ.2), !is.na(dep_df$PHQ.2))
t.test(!is.na(healthy_df$PHQ.9), !is.na(dep_df$PHQ.9))

#test first instance of phq2/9
df_unique_empi_healthy <- healthy_df %>% 
 group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup()

df_unique_empi_dep <- dep_df %>% 
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup()

t.test(!is.na(df_unique_empi_healthy$PHQ.2), !is.na(df_unique_empi_dep$PHQ.2))
t.test(!is.na(df_unique_empi_healthy$PHQ.9), !is.na(df_unique_empi_dep$PHQ.9))

print(paste0("Mean/SD depressed w/PHQ-2/9; healthy w/ph2/9 = ",
             round(mean(!is.na(df_unique_empi_dep$PHQ.2)),3),"[",round(sd(!is.na(df_unique_empi_dep$PHQ.2)),3), "];", 
             round(mean(!is.na(df_unique_empi_dep$PHQ.9)),3),
              "[",round(sd(!is.na(df_unique_empi_dep$PHQ.9)),3), "];",
             round(mean(!is.na(df_unique_empi_healthy$PHQ.2)),3),
             "[",round(sd(!is.na(df_unique_empi_healthy$PHQ.2)),3), "];",
             round(mean(!is.na(df_unique_empi_healthy$PHQ.9)),3),
             "[",round(sd(!is.na(df_unique_empi_healthy$PHQ.9)),3), "]"))

@@

@@

#Gender breakdown
#sex
hist(data_empi_acc_f_phq$sex_binarized[which(!is.na(data_empi_acc_f_phq$sex_binarized))], 
     main = paste0("Sex Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$sex_binarized))), "; M = ", length(which(data_empi_acc_f_phq$sex_binarized == 1)), "; F = ", length(which(data_empi_acc_f_phq$sex_binarized == 2)) ), 
     xlab = "Male = 1, Female = 2", 
     breaks = 2)

#sex with depression diagnosis
hist(dep_df$sex_binarized[which(!is.na(dep_df$sex_binarized))], 
     main = paste0("Sex Histo in ICD9/10 depression dx group : n = ", length(which(!is.na(dep_df$sex_binarized))), "; M = ", length(which(dep_df$sex_binarized == 1)), "; F = ", length(which(dep_df$sex_binarized == 2)) ), 
     xlab = "Male = 1, Female = 2", 
     breaks = 2)

#sex w/ phq2
hist(phq2_subset$sex_binarized[which(!is.na(phq2_subset$sex_binarized))], 
     main = paste0("Sex Histo w/phq2: n = ", length(which(!is.na(phq2_subset$sex_binarized))), "; M = ", length(which(phq2_subset$sex_binarized == 1)), "; F = ", length(which(phq2_subset$sex_binarized == 2)) ), 
     xlab = "Male = 1, Female = 2", 
     breaks = 2)

#sex w/phq9
hist(phq9_subset$sex_binarized[which(!is.na(phq9_subset$sex_binarized))], 
     main = paste0("Sex Histo w/phq9: n = ", length(which(!is.na(phq9_subset$sex_binarized))), "; M = ", length(which(phq9_subset$sex_binarized == 1)), "; F = ", length(which(phq9_subset$sex_binarized == 2)) ), 
     xlab = "Male = 1, Female = 2", 
     breaks = 2)

```{r frequencies_of_dep_in_phq2_vs_9}
num_dep_phq2 <- length(which(!is.na(data_empi_acc_f_phq$PHQ.2) & (data_empi_acc_f_phq$PHQ.2 >= 3))) #n = 130
num_dep_phq9 <- length(which(!is.na(data_empi_acc_f_phq$PHQ.9) & (data_empi_acc_f_phq$PHQ.9 >= 5))) #n = 184;
#https://www.sciencedirect.com/science/article/pii/B9780128179338000293 recommends using cut point of 11
num_total_phq2 <-length(which(!is.na(data_empi_acc_f_phq$PHQ.2)))
num_total_phq9 <-length(which(!is.na(data_empi_acc_f_phq$PHQ.9))) 
dep_total_phq_table <- data.frame(matrix(data = c(num_dep_phq2, num_dep_phq9, num_total_phq2, num_total_phq9), nrow = 2, ncol = 2))
names(dep_total_phq_table) <- c("dep", "healthy")
row.names(dep_total_phq_table) <- c("phq2", "phq9")
chisq_dep_phq2_v_phq9 <- chisq.test(dep_total_phq_table)
chisq_dep_phq2_v_phq9
print("observed")
chisq_dep_phq2_v_phq9$observed
print("expected")
chisq_dep_phq2_v_phq9$expected
print("residuals")
chisq_dep_phq2_v_phq9$residuals

```
```{r chi_sq_dep_versus_healthy_PHQ2_and_9}
num_dep_phq2 <- length(which(!is.na(dep_df$PHQ.2) & (dep_df$PHQ.2 >= 3))) #n = 83
num_dep_phq9 <- length(which(!is.na(dep_df$PHQ.9) & (dep_df$PHQ.9 >= 10))) #n = 79
num_healthy_phq2 <- length(which(!is.na(healthy_df$PHQ.2) & (healthy_df$PHQ.2 >= 3))) #n = 47
num_healthy_phq9 <- length(which(!is.na(healthy_df$PHQ.9) & (healthy_df$PHQ.9 >= 10))) #n = 32;
total_dep_phq2 <- length(which(!is.na(dep_df$PHQ.2))) #n = 83
total_dep_phq9 <- length(which(!is.na(dep_df$PHQ.9))) #n = 79
total_healthy_phq2 <- length(which(!is.na(healthy_df$PHQ.2))) #n = 47
total_healthy_phq9 <- length(which(!is.na(healthy_df$PHQ.9))) #n = 32;

#### phq2 symptomatic versus asymptomatic
phq2_table <- data.frame(matrix(data = c(num_dep_phq2,(total_dep_phq2-num_dep_phq2), num_healthy_phq2,   (total_healthy_phq2-num_healthy_phq2)), nrow = 2, ncol = 2))
names(phq2_table) <- c("dep", "healthy")
row.names(phq2_table) <- c("symptomatic", "asymptomatic")
chisq_dep_phq2 <- chisq.test(phq2_table)
chisq_dep_phq2
print("observed")
chisq_dep_phq2$observed
print("expected")
chisq_dep_phq2$expected
print("residuals")
chisq_dep_phq2$residuals

#### phq9 symptomatic versus asymptomatic
phq9_table <- data.frame(matrix(data = c(num_dep_phq9,(total_dep_phq9-num_dep_phq9), num_healthy_phq9,   (total_healthy_phq9 - num_healthy_phq9)), nrow = 2, ncol = 2))
names(phq9_table) <- c("dep", "healthy")
row.names(phq9_table) <- c("symptomatic", "asymptomatic")
chisq_dep_phq9 <- chisq.test(phq9_table)
chisq_dep_phq9
print("observed")
chisq_dep_phq9$observed
print("expected")
chisq_dep_phq9$expected
print("residuals")
chisq_dep_phq9$residuals

################### Unique

num_dep_phq2 <- length(which(!is.na(dep_df$PHQ.2) & (dep_df$PHQ.2 >= 3))) #n = 83
num_dep_phq9 <- length(which(!is.na(dep_df$PHQ.9) & (dep_df$PHQ.9 >= 10))) #n = 79
num_healthy_phq2 <- length(which(!is.na(healthy_df$PHQ.2) & (healthy_df$PHQ.2 >= 3))) #n = 47
num_healthy_phq9 <- length(which(!is.na(healthy_df$PHQ.9) & (healthy_df$PHQ.9 >= 10))) #n = 32;
total_dep_phq2 <- length(which(!is.na(dep_df$PHQ.2))) #n = 83
total_dep_phq9 <- length(which(!is.na(dep_df$PHQ.9))) #n = 79
total_healthy_phq2 <- length(which(!is.na(healthy_df$PHQ.2))) #n = 47
total_healthy_phq9 <- length(which(!is.na(healthy_df$PHQ.9))) #n = 32;

#patients who are depressed who score on phq2/9
num_dep_phq2_unique <- length(unique(dep_df$EMPI[which(dep_df$PHQ.2 >= 3)]))
num_dep_phq9_unique <- length(unique(dep_df$EMPI[which(dep_df$PHQ.9 >= 10)]))

#persons who are characterized as healthy but score on PHQ2/9
num_healthy_phq2_unique <- length(unique(healthy_df$EMPI[which(healthy_df$PHQ.2 >= 3)]))
num_healthy_phq9_unique <- length(unique(healthy_df$EMPI[which(healthy_df$PHQ.9 >= 10)]))

#total # of people who have phq2/9
asym_dep_phq2_unique <- length(unique(dep_df$EMPI[which(!is.na(dep_df$PHQ.2))])) - num_dep_phq2_unique
asym_dep_phq9_unique <- length(unique(dep_df$EMPI[which(!is.na(dep_df$PHQ.9))])) - num_dep_phq9_unique
asym_healthy_phq2_unique <- length(unique(healthy_df$EMPI[which(!is.na(healthy_df$PHQ.2))])) - num_healthy_phq2_unique
asym_healthy_phq9_unique <- length(unique(healthy_df$EMPI[which(!is.na(healthy_df$PHQ.9))])) - num_healthy_phq9_unique

#### phq2 symptomatic versus asymptomatic
phq2_table_unique <- data.frame(matrix(data = c(num_dep_phq2_unique,asym_dep_phq2_unique, num_healthy_phq2_unique, asym_healthy_phq2_unique), nrow = 2, ncol = 2))
names(phq2_table_unique) <- c("dep", "healthy")
row.names(phq2_table_unique) <- c("symptomatic", "asymptomatic")
chisq_dep_phq2_unique <- chisq.test(phq2_table_unique)
chisq_dep_phq2_unique
print("observed")
chisq_dep_phq2_unique$observed
print("expected")
chisq_dep_phq2_unique$expected
print("residuals")
chisq_dep_phq2_unique$residuals

#### phq9 symptomatic versus asymptomatic
phq9_table_unique <- data.frame(matrix(data = c(num_dep_phq9_unique,asym_dep_phq9_unique, num_healthy_phq9_unique,   asym_healthy_phq9_unique), nrow = 2, ncol = 2))
names(phq9_table_unique) <- c("dep", "healthy")
row.names(phq9_table_unique) <- c("symptomatic", "asymptomatic")
chisq_dep_phq9_unique <- chisq.test(phq9_table_unique)
chisq_dep_phq9_unique
print("observed")
chisq_dep_phq9_unique$observed
print("expected")
chisq_dep_phq9_unique$expected
print("residuals")
chisq_dep_phq9_unique$residuals

This next step, put together 9/29/2021, aims to define our control and case groups based on the following definitions-

Healthy group - No F33* code, and PHQ2 and 9 always 0

Lifetime depression - EITHER F33* diagnosis OR PHQ2 >= 3 OR PHQ9 >= 10
   - of note:
    *PHQ9 0-4 - no depression
    * 5-9 - mild depression
    * 10-14 - moderate depression
    * 15-19 - mod/severe
    * 20+ - severe

Logic
  1) First, separate out people with ICD10 code (F3*) into depressed and non depressed
       - dep_df is people with an F3* code
       - healthy df is people without
       - look for unique EMPIs in each of these groups

  2) Within "healthy" group, separate who has EVER had a lifetime history of testing positive on PHQ2 
        - Depression is PHQ2 >=3 at any point
        - Healthy if PHQ2 = 0
        - Look for unique EMPIs in these groups

  3) Within "healthy" group, separate who has EVER had a lifetime history of testing positive on PHQ9
        - Depression is PHQ9 >= 10 at any point
        - Healthy if PHQ9 = 0
        - Look for unique EMPIs in these groups

  4) Combine for depressed group and healthy
        - Look for people with PHQ2 or 9 meeting criteria for depression that aren't in depression group, and add them -> dep_num_icd10_plus_phq2_and_phq9
        - remove people from healthy_df if people had EMPIs in phq2 or 9 depressed group -> healthy_num_no_icd10_plus_phq2_and_phq9
        - make separate group that only includes healthies for whom we have phq2 or 9 -> healthy_num_no_icd10_plus_phq2_and_phq9_MUST_HAVE_PHQ2_OR_9

```{r defining_control_and_dep_groups}

##########################################################
#### Step 1- make F3* depressed and healthy groups* ######
##########################################################

########
### separate out depressed group
########
print("separate out depressed group")
dep_df <- data_empi_acc_f_phq[grep("F3", data_empi_acc_f_phq$ICD10),] #n = 2123
num_depression <- dim(dep_df)[1]
num_unique_EMPI_and_dep <- length(unique(dep_df$EMPI)) #n = 493
print(paste0("N with ICD 9/10 codes for depression = ", num_depression))
print(paste0("N with ICD 9/10 codes for depression and unique EMPI in whole depression group = ", num_unique_EMPI_and_dep, " out of ", num_depression))

#######
### separate out healthy group
#######
print("separate out healthy group")
healthy_df <- data_empi_acc_f_phq[-grep("F3", data_empi_acc_f_phq$ICD10),] #n = 14707
num_healthy <- dim(healthy_df)[1] 
num_unique_EMPI_and_healthy <- length(unique(healthy_df$EMPI)) #n = 3244
print(paste0("N with WITHOUT depression ICD 9/10 codes (i.e. \"healthy\"): ", num_healthy))
print(paste0("N with WITHOUT ICD 9/10 depression codes for depression and unique EMPI = ", num_unique_EMPI_and_healthy, " out of ", num_healthy))

##########################################################
#### Step 2- Identify PHQ2 >=3 in healthy group     ######
##########################################################

print("Get people in healthy group with phq2s for dep/healthy")
# num in healthy group with PHQ2
phq2_subset <- healthy_df[which(!is.na(healthy_df$PHQ.2)),] #n=3646, unique 698
depressed_phq2_subset <- phq2_subset[which(phq2_subset$PHQ.2 >= 3),] #n = 47 
depressed_unique_phq2_empis <- unique(depressed_phq2_subset$EMPI) #n=10
healthy_phq2_subset <- phq2_subset[which(phq2_subset$PHQ.2 == 0),] #n = 3326
healthy_unique_phq2_empis <- unique(healthy_phq2_subset$EMPI) #n = 631

print(paste0("PHQ2 breakdown in healthy group (total n with score = ", dim(phq2_subset)[1], "[unique = ", length(unique(phq2_subset$EMPI)),"]): PHQ2 >= 3 :", dim(depressed_phq2_subset)[1], "[unique = ", length(depressed_unique_phq2_empis), "]; PHQ2 == 0 : ", dim(healthy_phq2_subset)[1], "[unique = ", length(healthy_unique_phq2_empis),"]"))

##########################################################
#### Step 3- Identify PHQ9 >=10 in healthy group    ######
##########################################################

print("Get people in healthy group with phq9s for dep/healthy")
# num in healthy group with PHQ9
phq9_subset <- healthy_df[which(!is.na(healthy_df$PHQ.9)),] #n=121
depressed_phq9_subset <- phq9_subset[which(phq9_subset$PHQ.9 >= 10),] #n = 32 
depressed_unique_phq9_empis <- unique(depressed_phq9_subset$EMPI) #n=9
healthy_phq9_subset <- phq9_subset[which(phq9_subset$PHQ.9 == 0),] #n = 21
healthy_unique_phq9_empis <- unique(healthy_phq9_subset$EMPI) #n =4

print(paste0("PHQ9 breakdown in healthy group (total n with score = ", dim(phq9_subset)[1], "[unique = ", length(unique(phq9_subset$EMPI)),"]): PHQ9 >= 10 :", dim(depressed_phq9_subset)[1], "[unique = ", length(depressed_unique_phq9_empis), "]; PHQ9 == 0 : ", dim(healthy_phq9_subset)[1], "[unique = ", length(unique(healthy_phq9_subset$EMPI)),"]"))

##########################################################
#Step 4 - Combine for depressed group and healthy groups #
##########################################################

print("Combine groups for new cohort")
gain_in_dep_group <- length(depressed_unique_phq2_empis) + length(depressed_unique_phq9_empis)
dep_num_icd10_plus_phq2_and_phq9 <- num_unique_EMPI_and_dep + gain_in_dep_group # we can do this because we know they are each exclusive - num_unique_empi_and_dep from icd10 group, and no overlap in phq2 and 9 group
print(paste0("number of UNIQUE people with an icd10 code for Depression OR scored positive on PHQ2(>=3) or PHQ9(>=10) : ", dep_num_icd10_plus_phq2_and_phq9, ", for a gain of n = ", gain_in_dep_group))

healthy_num_no_icd10_plus_phq2_and_phq9 <- num_unique_EMPI_and_healthy - gain_in_dep_group #take number of people who don't have depression diagnosis and don't have a depression phq2/9

print(paste0("number of UNIQUE people with NO icd10 code for Depression and NEVER had a PHQ2(>=3) or PHQ9(>=10) : ", healthy_num_no_icd10_plus_phq2_and_phq9, ", for a loss of n = ", gain_in_dep_group))

healthy_num_no_icd10_plus_phq2_and_phq9_MUST_HAVE_PHQ2_OR_9 <- length(healthy_unique_phq2_empis) + length(healthy_unique_phq9_empis)
print(paste0("number of UNIQUE people with NO icd10 code for Depression and HAS HAD AT LEAST 1 PHQ2 or 9 that was 0) : ", healthy_num_no_icd10_plus_phq2_and_phq9_MUST_HAVE_PHQ2_OR_9))

print(paste0("Cleanest cohort (depressed in icd9/10 or Phq2/9; healthy by at least one phq2/9 and never depression dx ---->>>>  Depressed: ", dep_num_icd10_plus_phq2_and_phq9, "; Healthy: ", healthy_num_no_icd10_plus_phq2_and_phq9_MUST_HAVE_PHQ2_OR_9))

##### THESE ARE THE DATAFRAMES I WANT TO USE GOING FORWARD ###
#all empis for depressed group include people with depresssion dx, and people who got in for phq2 or phq9
empis_for_depressed_group <- append(
  append(unique(dep_df$EMPI), depressed_unique_phq2_empis), depressed_unique_phq9_empis)

final_depressed_group_withICD_AND_depressed_phq2_or_9 <- 
  data_empi_acc_f_phq %>%
filter(EMPI %in% empis_for_depressed_group)

#------
empis_for_healthy_group <- append(healthy_unique_phq2_empis, healthy_unique_phq9_empis)

final_healthy_group_withNOICD_AND_phq2_or_9_0<- 
  data_empi_acc_f_phq %>%
filter(EMPI %in% empis_for_healthy_group)

#------
### add a column to the original dataframe with a 0 if not in any group, -1 if healthy, and 1 if depressed
data_empi_acc_f_phq$depGroupVar = 0
data_empi_acc_f_phq$depGroupVar[which(data_empi_acc_f_phq$EMPI %in% empis_for_depressed_group)] = 1
data_empi_acc_f_phq$depGroupVar[which(data_empi_acc_f_phq$EMPI %in% empis_for_healthy_group)] = -1
data_empi_acc_f_phq$depGroupVar = as.factor(data_empi_acc_f_phq$depGroupVar)

@@

#Gender breakdown
#sex
hist(data_empi_acc_f_phq$sex_binarized[which(!is.na(data_empi_acc_f_phq$sex_binarized))], 
     main = paste0("Sex Histo : n = ", length(which(!is.na(data_empi_acc_f_phq$sex_binarized))), "; M = ", length(which(data_empi_acc_f_phq$sex_binarized == 1)), "; F = ", length(which(data_empi_acc_f_phq$sex_binarized == 2)) ), 
     xlab = "Male = 1, Female = 2", 
     breaks = 2)

```{r comparing_out_group_with_Tim_group}
##### checking to see whether we have clinical data for patients who have already been segmented as of 5/4/2021
### had to do some parsing of tim's spreadsheet: /Users/eballer/BBL/msdepression/data/insurance_images_for_depression.csv
### used the following commands to extract the empi and the date -> in tim's they are in the format sub-***** and ses-YYYYMMDD

@@

#segmented <- read.csv("/Users/eballer/BBL/msdepression/data/insurance_images_for_depression_with_accession.csv", header = F, sep = ",") #n = 4468
segmented <- read.csv("/Users/eballer/BBL/msdepression/data/insurance_images_for_depression3.csv", header = F, sep = ",") #n = 2833

names(segmented) <- c("EMPI", "EXAM_DATE", "SES", "ACCESSION_NUM")

@@

print(paste0("Num people in previously curated group and Erica's group -> ", dim(overlap_group)[1]))

@@

print(paste0("N with ICD 9/10 codes for depression and unique EMPI = ", num_unique_EMPI_and_dep, " out of total depressed - ", dim(dep_df)[1]))

@@

```{r depression_vs_healthy_melissa_qc_group}

### this chunk can be run separately from the rest of the analysis
### A lot of it is repeat, but I wanted it to be independent
### Started 20211119

homedir <- "/Users/eballer/BBL/msdepression/"
source("~/BBL/msdepression/scripts/msdepression_functions.R")

##########################################################
####                   Prepare the date               ####
##########################################################
#load our df

#from Jan 2021 pull
data <- read.csv("/Users/eballer/BBL/msdepression/data/dac/investigatingdepressioninmspatients_dates_right_format.csv", header = TRUE, stringsAsFactors = FALSE, na.strings = c("NULL"), sep =",")

#make accession # and lab findings into integer type, contained in row 28:42, Margin=2 to work on columns
data_empi_acc_f_phq$ACCESSION_NUM <- as.integer(data_empi_acc_f_phq$ACCESSION_NUM)
data_empi_acc_f_phq[27:29] <- apply(X = data_empi_acc_f_phq[27:29], FUN = as.integer, MARGIN = 2)
data_empi_acc_f_phq[31:45] <- apply(X = data_empi_acc_f_phq[31:45], FUN = as.integer, MARGIN = 2)
data_empi_acc_f_phq$sex_binarized <- ifelse(data_empi_acc_f_phq$SEX == "MALE", 1, 2)
data_empi_acc_f_phq$osex <- ordered(data_empi_acc_f_phq$sex_binarized)
data_empi_acc_f_phq$race_binarized <- ifelse(data_empi_acc_f_phq$RACE == "WHITE", 1, 2)
#get date into better format
data_empi_acc_f_phq$EXAM_DATE <- as.Date(data_empi_acc_f_phq$BEGIN_EXAM_DTTM, format = "%m/%d/%y")
data_empi_acc_f_phq$EXAM_DATE <- as.numeric(gsub(data_empi_acc_f_phq$EXAM_DATE, pattern = "-", replacement = ""))

##########################################################
#### Step 1- make F3* depressed and healthy groups* ######
##########################################################
  ### This is identical to up above, but repeated so I can run this chunk separately

########
### separate out depressed group
########
print("separate out depressed group")
dep_df <- data_empi_acc_f_phq[grep("F3", data_empi_acc_f_phq$ICD10),] #n = 2123
num_depression <- dim(dep_df)[1]
num_unique_EMPI_and_dep <- length(unique(dep_df$EMPI)) #n = 493
print(paste0("N with ICD 9/10 codes for depression = ", num_depression))
print(paste0("N with ICD 9/10 codes for depression and unique EMPI in whole depression group = ", num_unique_EMPI_and_dep, " out of ", num_depression))

#find number of participants who meet criteria for depression in the PHQ2/9 symptom rating
print(paste0("N with ICD 9/10 codes for depression, PHQ2 >= 3, unique = ", length(!is.na(unique(dep_df$EMPI[dep_df$PHQ.2 >= 3])))))
print(paste0("N with ICD 9/10 codes for depression, PHQ9 >= 10, unique = ", length(!is.na(unique(dep_df$EMPI[dep_df$PHQ.9 >= 10])))))

#######
### separate out healthy group
#######
print("separate out healthy group")
healthy_df <- data_empi_acc_f_phq[-grep("F3", data_empi_acc_f_phq$ICD10),] #n = 14707
num_healthy <- dim(healthy_df)[1] 
num_unique_EMPI_and_healthy <- length(unique(healthy_df$EMPI)) #n = 3244
print(paste0("N with WITHOUT depression ICD 9/10 codes (i.e. \"healthy\"): ", num_healthy))
print(paste0("N with WITHOUT ICD 9/10 depression codes for depression and unique EMPI = ", num_unique_EMPI_and_healthy, " out of ", num_healthy))

##########################################################
#### Step 2- Identify PHQ2 >=3 in healthy group     ######
##########################################################

print("Get people in healthy group with phq2s for dep/healthy")
# num in healthy group with PHQ2
phq2_subset <- healthy_df[which(!is.na(healthy_df$PHQ.2)),] #n=3646, unique 698
depressed_phq2_subset <- phq2_subset[which(phq2_subset$PHQ.2 >= 3),] #n = 47 
depressed_unique_phq2_empis <- unique(depressed_phq2_subset$EMPI) #n=10
healthy_phq2_subset <- phq2_subset[which(phq2_subset$PHQ.2 == 0),] #n = 3326
healthy_unique_phq2_empis <- unique(healthy_phq2_subset$EMPI) #n = 631

print(paste0("PHQ2 breakdown in healthy group (total n with score = ", dim(phq2_subset)[1], "[unique = ", length(unique(phq2_subset$EMPI)),"]): PHQ2 >= 3 :", dim(depressed_phq2_subset)[1], "[unique = ", length(depressed_unique_phq2_empis), "]; PHQ2 == 0 : ", dim(healthy_phq2_subset)[1], "[unique = ", length(healthy_unique_phq2_empis),"]"))

##########################################################
#### Step 3- Identify PHQ9 >=10 in healthy group    ######
##########################################################

print("Get people in healthy group with phq9s for dep/healthy")
# num in healthy group with PHQ9
phq9_subset <- healthy_df[which(!is.na(healthy_df$PHQ.9)),] #n=121
depressed_phq9_subset <- phq9_subset[which(phq9_subset$PHQ.9 >= 10),] #n = 32 
depressed_unique_phq9_empis <- unique(depressed_phq9_subset$EMPI) #n=9
healthy_phq9_subset <- phq9_subset[which(phq9_subset$PHQ.9 == 0),] #n = 21
healthy_unique_phq9_empis <- unique(healthy_phq9_subset$EMPI) #n =4

print(paste0("PHQ9 breakdown in healthy group (total n with score = ", dim(phq9_subset)[1], "[unique = ", length(unique(phq9_subset$EMPI)),"]): PHQ9 >= 10 :", dim(depressed_phq9_subset)[1], "[unique = ", length(depressed_unique_phq9_empis), "]; PHQ9 == 0 : ", dim(healthy_phq9_subset)[1], "[unique = ", length(unique(healthy_phq9_subset$EMPI)),"]"))

##########################################################
#Step 4 - Combine for depressed group and healthy groups #
##########################################################

print("Combine groups for new cohort")
gain_in_dep_group <- length(depressed_unique_phq2_empis) + length(depressed_unique_phq9_empis)
dep_num_icd10_plus_phq2_and_phq9 <- num_unique_EMPI_and_dep + gain_in_dep_group # we can do this because we know they are each exclusive - num_unique_empi_and_dep from icd10 group, and no overlap in phq2 and 9 group
print(paste0("number of UNIQUE people with an icd10 code for Depression OR scored positive on PHQ2(>=3) or PHQ9(>=10) : ", dep_num_icd10_plus_phq2_and_phq9, ", for a gain of n = ", gain_in_dep_group))

healthy_num_no_icd10_plus_phq2_and_phq9 <- num_unique_EMPI_and_healthy - gain_in_dep_group #take number of people who don't have depression diagnosis and don't have a depression phq2/9

print(paste0("number of people with NO icd10 code for Depression and NEVER had a PHQ2(>=3) or PHQ9(>=10) : ", healthy_num_no_icd10_plus_phq2_and_phq9, ", for a loss of n = ", gain_in_dep_group))

healthy_num_no_icd10_plus_phq2_and_phq9_MUST_HAVE_PHQ2_OR_9 <- length(healthy_unique_phq2_empis) + length(healthy_unique_phq9_empis)
print(paste0("number of UNIQUE people with NO icd10 code for Depression and HAS HAD AT LEAST 1 PHQ2 or 9 that was 0) : ", healthy_num_no_icd10_plus_phq2_and_phq9_MUST_HAVE_PHQ2_OR_9))

print(paste0("Cleanest cohort (depressed in icd9/10 or Phq2/9; healthy by at least one phq2/9 and never depression dx UNIQUE ---->>>>  Depressed: ", dep_num_icd10_plus_phq2_and_phq9, "; Healthy: ", healthy_num_no_icd10_plus_phq2_and_phq9_MUST_HAVE_PHQ2_OR_9))

##### THESE ARE THE DATAFRAMES I WANT TO USE GOING FORWARD ###
#all empis for depressed group include people with depresssion dx, and people who got in for phq2 or phq9
empis_for_depressed_group <- append(
  append(unique(dep_df$EMPI), depressed_unique_phq2_empis), depressed_unique_phq9_empis)

final_depressed_group_withICD_AND_depressed_phq2_or_9 <- 
  data_empi_acc_f_phq %>%
filter(EMPI %in% empis_for_depressed_group)

#------
empis_for_healthy_group <- append(healthy_unique_phq2_empis, healthy_unique_phq9_empis)
